---
title: "Medicare data challenge"
author: "Ying Li"
date: "2017.11.27"
output:
  html_document: default
  pdf_document: default
subtiltle: An exploration of home health care agencies
---

```{r setup, include=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#1. Why home health care?

The reason I learn statistics is that I am obsessed with health-related statistics. After I came to the US when I was checking my health insurance, I find that there are lots of home health care agencies that can bring health care services to your home. This is very unusual compared to China and I am always interested in things that are new to me. So I decided to focus on home health care agncies.

When people are choosing their own home health care agency, there are many things that take into consideration, like:

1. What kind of health care service are they offering?

2. Are the health provider nice?

3. What are the prices?

4. Are the doctors experienced?


What's more? A lot of people write reviews for certain agncy, What factors could influence them when writing a reviews? Why would they give an agency five stars?

For this project, I will be finding out the factors that could influence the star rate of each agency start. And does the state location influence the star rating for each agency?


#2. Subsetting the Large Datasets

I use the website "https://data.medicare.gov/Home-Health-Compare/Home-Health-Care-Agencies/6jpm-sxkc" to get the data. This datasets consists of lots of variables including the address of each agency and some footnote for the variables. So I delete those variables that I think is not relevant to my project. From the data I subsetted the data by keeping data with state category and some other variables that I think that could influence that star rating of an agency.

After subsetting the datasets, my final datasets have 11742 agencies and their ratings.

The column names of this datasets are all long senteces therefore I make some abbreviation to simplfy the column names.


```{r,warning=FALSE}
library(ggplot2)
library(dplyr)
hca1<-read.csv("Home_Health_Care_Agencies.csv")
colnames(hca1)[1]<-c("State")
hca2<-hca1[-c(2,3,4,6,7,8,15,17,19,21,23,25,27,29,31,33,35,37,39,41,43,45,47,49,50,51,52,53,54)]
colnames(hca2)<-c("state","city","NC","PT","OT","SP","MS","HHA","rating","begin_care","teach_drug","falling","depression","flu_shot","pneu_vac","foot_care","walking","bed","bathing","pain","breath","operation","take_drug","hospital","ER")
hca2$NC<-ifelse(hca2$NC=="true",1,0)
hca2$PT<-ifelse(hca2$PT=="true",1,0)
hca2$OT<-ifelse(hca2$OT=="true",1,0)
hca2$SP<-ifelse(hca2$SP=="true",1,0)
hca2$MS<-ifelse(hca2$MS=="true",1,0)
hca2$HHA<-ifelse(hca2$HHA=="true",1,0)
hca<-hca2[-3]
hca<-na.omit(hca)
```

#3. EDA

##3.1 counts of the star ratings

```{r,warning=FALSE}
hist(hca$rating,main = "counts of star ratings",xlab="star rating",col="pink",border = "blue")
```

Based on the histogram, we can conclude that the rating at four stars has the most counts, then follows two stars and three stars. Nearlly no agency have only one star, but there are some agencies that have five stars. Since this datastes have state category, so my next plot should be the counts of star ratings for each state. However there is fifty states in the US, so I decided to plot the mean star ratings for each state.

##3.2 star ratings for each states

```{r,warning=FALSE}
hca_state<-hca[,c(1,8)]
hca_state<-na.omit(hca_state)
ggplot(hca_state,aes(x=state,y=rating))+geom_boxplot()+labs(title="boxplot fo the rating of each group",x="state",y="rating")+theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Based on the boxplot, I find that most of the states have 3.5 star rating as mean and the lowest star rating is around two and the highest star rating is around four. Since I have built plots to see the influence of the state category, so I decided that I should include maybe random slope or intercept in my model. But for my next step, I will be looking at some other factors that I think might have some influence on the star rating.

##3.3 EDA for some states 

```{r}
hca_4state<-filter(hca_state,state==c("MA","NY","AL"))
hca_4state<-na.omit(hca_4state)
ggplot(hca_4state,aes(state))+geom_bar(aes(fill=factor(rating)))
```

I only choose three typical states to see whether there are differences among different states. Based on the plot, we can see that the raings for differenct states are quit uniqe. For state AL, most of the rating are around 4 stars, while for MA, it is 3.5 stars and for NY, it is 3. Therefore we can see that there are differences among different states.

##3.4 EDA of some influencing factors

From my own experience, I choosed some factors that I think are not neglectable. And I will be plot a graph hopinng it will show the relationship of the star rating and the factors.
```{r,warning=FALSE}
ggplot(hca,aes(x=flu_shot,y=rating))+geom_jitter()+geom_smooth(method = "lm")
ggplot(hca,aes(x=operation,y=rating))+geom_jitter()+geom_smooth(method = "lm")
ggplot(hca,aes(x=hospital,y=rating))+geom_jitter()+geom_smooth(method = "lm")
ggplot(hca,aes(x=HHA,y=rating))+geom_jitter()
```

Based on these three plots, it looks like that the first two influencing factors might all hava a positive impact on star rating. For the first plot, we can see that most of the points are jittering around 75, and the star rating is jittering around 4 stars. This is the same for the other two variables. For the last influencing factor, we can see that there is a negative impact on star raing, which in fact makes sense. But for the variables: whether they offer home health aid servece, we can see that most of the agencies offer home health aid. And there is apparently a great difference in star rating among agencies that offer home health aid and agencies that do not offer home health aid. For my next step I will be fitting a few mutilevel models on star rating using these factors that I just plotted.

#4. Multilevel Modeling

##4.1 random intercept

I want to predict star rating based on the flu shot duration, and how often the patients had to be admitted to the hospital, how often patients' wounds improved or healed after an operation and whether they offer home health aid service and whether they offer occupational therapy service.

To find the good model fit, I will be playing around with the dataset first and fitting lots of models to try to find the great model fit.

```{r,warning=FALSE}
library(lme4)
fit1<-lmer(rating~flu_shot+operation+hospital+(1|state),data=hca)
summary(fit1)
fit2<-lmer(rating~flu_shot+operation+hospital+HHA+(1|state),data=hca)
summary(fit2)
AIC(fit1,fit2)
fit3<-lmer(rating~flu_shot*state+operation+hospital+HHA+(1|state),data=hca)
fit4<-lmer(rating~flu_shot+operation*state+hospital+HHA+(1|state),data=hca)
fit5<-lmer(rating~flu_shot+operation+hospital*state+HHA+(1|state),data=hca)
fit6<-lmer(rating~flu_shot+operation+hospital+HHA*state+(1|state),data=hca)
AIC(fit2,fit3,fit4,fit5,fit6)
```

Based on the AIC result, the best model fit is model fit2, including factors such as HHA, flu_shot, operation, hospital and the group variable state.

##4.2 random slope

For the previous models that I fit, there are only random intercept effect. I picked the best model fit and now I will be fitting random slope models.

```{r,warning=FALSE}
fit7<-lmer(rating~flu_shot+operation+hospital+HHA+(1+hospital|state),data=hca)
fit8<-lmer(rating~flu_shot+operation+hospital+HHA+(1+HHA|state),data=hca)
fit9<-lmer(rating~flu_shot+operation+hospital+HHA+(1+operation|state),data=hca)
fit10<-lmer(rating~flu_shot+operation+hospital+HHA+(1+flu_shot|state),data=hca)
AIC(fit7,fit8,fit9,fit10)
fit11<-lmer(rating~flu_shot+operation+hospital*state+HHA+(1+flu_shot|state),data=hca)
fit12<-lmer(rating~flu_shot+operation*state+hospital+HHA+(1+flu_shot|state),data=hca)
fit13<-lmer(rating~flu_shot*state+operation+hospital+HHA+(1+flu_shot|state),data=hca)
fit14<-lmer(rating~flu_shot+operation+hospital+HHA*state+(1+flu_shot|state),data=hca)
AIC(fit10,fit11,fit12,fit13,fit14)

```

Based on the AIC result, the best model fit is model fit10, including factors such as HHA, flu_shot, operation, hospital and random slope group effect on flu shot.

Now, I have two good-fit models which are random intercept effect model and random slope effect model. For my next step, I will be doing some model checking.

##4.3 Model checking

For this model checking part, I will be using likelihood ratio test and some plots to check my models.

```{r,warning=FALSE}
library(arm)
plot(fit2)
plot(fit10)
binnedplot(predict(fit2),resid(fit2))
binnedplot(predict(fit10),resid(fit10))
anova(fit2,fit10)
summary(fit10)
```

From the residual plot, we can see that most of the dots are jitering around the 3 and 4. It makes sense because most of the star rating are around 4. From the binned residual plot, we can see that for the two models, most of the points fall within the boundary. At last, comparing the two model with AIC, BIC and deviance, there is only a slight difference in AIC, BIC and deviance, however for the model fit10, the model is nearly unidentifiable because the eigenvalue is very large.
Based on this, I assume the best model fit is model fit2.

Overall, I fit lots of random intercept models and random slope models and find the best fit for these two kinds, and then do some model checking and find the best model fit model fit2 which is a random intercept model.

For my next step, I will be doing some interpretation of model fit2.

##4.4 interpretation of the model

```{r,warning=FALSE}
summary(fit2)
library(merTools)
plotFEsim(FEsim(fit2))

```

From this plot, we could see that the HHA which mean whether the agency offers home health aid or not have negative effect on star rating. And hospital which means how often home health patients who have had a recent hospital stay had to be readmitted to the hospital has a negaive effect on the star rating. While flu shot and operation which means how often the home health team determined whether patients received a flu shot for the curent flu season and how often patients wounds improved or healed after an operation have an positive effect on the star rating. As for the group factor state. From my model it is a random intercept model and the distribution of the state intercept is 0.1.

#5. Further discussion

For my home health care agency exploration, I fit several multilevel models and find the best model fit. However, I only use some certain variables as factors, but there are many other variables in the dataset that can be used to determine the influence on star rating of each agency. Futhermore, there might be some correlation between the variables in the models that I fit. So for my future direction, I will be focusing on some other variables and the correlation beween the variables.




